---
title: "tidysoiltexture — Grid Expansion Tests"
subtitle: "Systematic validation of `classify_texture` and `gg_texture_triangle` across all 12 USDA texture classes"
author: "TidySoil"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    theme: flatly
    df-print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Setup

```{r packages}
# Load from local source (not yet installed to library)
pkg_root <- rprojroot::find_root(rprojroot::is_rstudio_project)
devtools::load_all(file.path(pkg_root, "tidysoiltexture"), quiet = TRUE)

library(tidyverse)
library(knitr)
```

### Reference: USDA class centroids

One representative point per texture class, chosen to lie clearly within its
polygon (not near any boundary).

```{r centroids}
class_centroids <- tribble(
  ~expected_class,    ~expected_abbr, ~sand, ~silt, ~clay,
  "sand",             "Sa",            92,     5,     3,
  "loamy sand",       "LoSa",          80,    12,     8,
  "sandy loam",       "SaLo",          65,    22,    13,
  "loam",             "Lo",            41,    39,    20,
  "silty loam",       "SiLo",          20,    65,    15,
  "silt",             "Si",             7,    85,     8,
  "sandy clay loam",  "SaClLo",        60,    13,    27,
  "clay loam",        "ClLo",          32,    34,    34,
  "silty clay loam",  "SiClLo",        10,    56,    34,
  "sandy clay",       "SaCl",          52,     6,    42,
  "silty clay",       "SiCl",           6,    47,    47,
  "clay",             "Cl",            20,    20,    60
)

kable(class_centroids,
      caption = "Representative centroids used as reference grid")
```

---

## Part 1 — `classify_texture()`

### 1.1 Centroid grid: all 12 classes

```{r classify-centroids}
centroid_results <- class_centroids |>
  classify_texture(sand = sand, silt = silt, clay = clay)

kable(
  select(centroid_results, expected_class, .texture_class, .texture_abbr,
         sand, silt, clay),
  caption = "Classification of all 12 centroids"
)
```

```{r classify-centroid-check}
centroid_check <- centroid_results |>
  mutate(
    class_pass = expected_class == .texture_class,
    abbr_pass  = expected_abbr  == .texture_abbr
  )

all_pass <- all(centroid_check$class_pass) && all(centroid_check$abbr_pass)
cat(if (all_pass) "✓ All 12 centroids classified correctly.\n"
    else "✗ Some centroids misclassified — see table above.\n")

stopifnot("Centroid misclassified" = all_pass)
```

### 1.2 Multi-sample grid: 5 points per class

For each of the 12 classes we define 5 spread-out interior samples.
This gives a **60-row** test grid.

```{r multi-sample-grid}
class_samples <- tribble(
  ~expected_class,   ~sand, ~silt, ~clay,
  # sand  (>85% sand, silt+1.5*clay < 15)
  "sand",            90,    7,    3,
  "sand",            95,    3,    2,
  "sand",            88,    8,    4,
  "sand",            93,    5,    2,
  "sand",            97,    2,    1,
  # loamy sand
  "loamy sand",      83,    10,   7,
  "loamy sand",      78,    15,   7,
  "loamy sand",      80,    12,   8,
  "loamy sand",      82,    8,   10,
  "loamy sand",      86,    6,    8,
  # sandy loam
  "sandy loam",      70,    18,  12,
  "sandy loam",      60,    27,  13,
  "sandy loam",      75,    12,  13,
  "sandy loam",      55,    30,  15,
  "sandy loam",      65,    20,  15,
  # loam
  "loam",            43,    38,  19,
  "loam",            35,    42,  23,
  "loam",            48,    30,  22,
  "loam",            40,    37,  23,
  "loam",            30,    48,  22,
  # silty loam  (labelled as the package returns it)
  "silty loam",      20,    60,  20,
  "silty loam",      15,    70,  15,
  "silty loam",      25,    55,  20,
  "silty loam",      10,    75,  15,
  "silty loam",      28,    58,  14,
  # silt
  "silt",            5,     88,   7,
  "silt",            3,     92,   5,
  "silt",            10,    82,   8,
  "silt",            7,     86,   7,
  "silt",            2,     94,   4,
  # sandy clay loam
  "sandy clay loam", 58,    14,  28,
  "sandy clay loam", 63,    10,  27,
  "sandy clay loam", 55,    17,  28,
  "sandy clay loam", 68,     7,  25,
  "sandy clay loam", 50,    22,  28,
  # clay loam
  "clay loam",       33,    35,  32,
  "clay loam",       28,    38,  34,
  "clay loam",       38,    30,  32,
  "clay loam",       25,    40,  35,
  "clay loam",       42,    27,  31,
  # silty clay loam
  "silty clay loam", 12,    55,  33,
  "silty clay loam", 8,     62,  30,
  "silty clay loam", 15,    52,  33,
  "silty clay loam", 5,     65,  30,
  "silty clay loam", 15,    52,  33,
  # sandy clay
  "sandy clay",      55,     4,  41,
  "sandy clay",      50,     8,  42,
  "sandy clay",      60,     3,  37,
  "sandy clay",      48,     5,  47,
  "sandy clay",      53,     7,  40,
  # silty clay
  "silty clay",      5,     50,  45,
  "silty clay",      3,     55,  42,
  "silty clay",      8,     45,  47,
  "silty clay",      2,     58,  40,
  "silty clay",      10,    43,  47,
  # clay
  "clay",            22,    18,  60,
  "clay",            15,    25,  60,
  "clay",            28,    14,  58,
  "clay",            10,    30,  60,
  "clay",            35,    10,  55
)

cat("Grid dimensions:", nrow(class_samples), "rows ×",
    ncol(class_samples), "columns\n")
```

```{r multi-sample-classify}
sample_results <- class_samples |>
  classify_texture(sand = sand, silt = silt, clay = clay)

kable(
  sample_results |>
    count(expected_class, .texture_class) |>
    pivot_wider(names_from = .texture_class,
                values_from = n, values_fill = 0L),
  caption = "Confusion matrix: expected (rows) vs predicted (columns)"
)
```

```{r multi-sample-accuracy}
accuracy_by_class <- sample_results |>
  mutate(correct = expected_class == .texture_class) |>
  group_by(expected_class) |>
  summarise(
    n_total   = n(),
    n_correct = sum(correct),
    accuracy  = n_correct / n_total,
    .groups   = "drop"
  )

kable(accuracy_by_class,
      caption = "Per-class accuracy on the 60-point grid")

overall_acc <- mean(sample_results$expected_class == sample_results$.texture_class)
cat(sprintf("Overall accuracy: %.1f%% (%d / %d samples)\n",
            overall_acc * 100,
            sum(sample_results$expected_class == sample_results$.texture_class),
            nrow(sample_results)))
```

### 1.3 Column name vs scalar equivalence

For every centroid, classify via columns and via scalars; results must match.

```{r classify-equiv}
equiv_check <- class_centroids |>
  rowwise() |>
  mutate(
    class_col    = classify_texture(
      tibble(sand = sand, silt = silt, clay = clay),
      sand = sand, silt = silt, clay = clay)$.texture_class,
    class_scalar = classify_texture(
      tibble(id = 1L),
      sand = sand, silt = silt, clay = clay)$.texture_class,
    pass = class_col == class_scalar
  ) |>
  ungroup() |>
  select(expected_class, class_col, class_scalar, pass)

kable(equiv_check,
      caption = "Column-name vs scalar equivalence across all classes")

stopifnot("Column vs scalar mismatch" = all(equiv_check$pass))
cat("✓ Column/scalar equivalence confirmed for all 12 classes.\n")
```

### 1.4 Output structure

```{r classify-structure}
result <- classify_texture(
  tibble(sand = 40, silt = 40, clay = 20, site = "A"),
  sand = sand, silt = silt, clay = clay
)

stopifnot(
  "Not a tibble"            = inherits(result, "tbl_df"),
  "Missing .texture_class"  = ".texture_class" %in% names(result),
  "Missing .texture_abbr"   = ".texture_abbr"  %in% names(result),
  "Original cols preserved" = "site"           %in% names(result),
  "Row count changed"       = nrow(result) == 1L
)
cat("✓ Output structure is correct (tibble, both class columns, original cols preserved).\n")
```

### 1.5 Input validation

```{r classify-validation}
# sand + silt + clay ≠ 100
tryCatch(
  classify_texture(tibble(sand = 50, silt = 30, clay = 10),
                   sand = sand, silt = silt, clay = clay),
  error = \(e) cat("✓ Sum ≠ 100:", conditionMessage(e), "\n")
)

# unsupported system
tryCatch(
  classify_texture(tibble(sand = 40, silt = 40, clay = 20),
                   sand = sand, silt = silt, clay = clay, system = "EU"),
  error = \(e) cat("✓ Bad system:", conditionMessage(e), "\n")
)
```

### 1.6 Expand-grid: systematic ternary sweep

We create a systematic grid over the ternary diagram (every 5 %) and verify
that every valid point — including axis edges and triangle vertices — receives
a non-NA classification. The `point_in_polygon()` implementation uses an
on-edge collinearity check before ray-casting, so boundary points are handled
correctly.

```{r ternary-sweep}
sweep_grid <- expand_grid(
  sand = seq(0, 100, by = 5),
  silt = seq(0, 100, by = 5)
) |>
  mutate(clay = 100 - sand - silt) |>
  filter(clay >= 0, sand + silt + clay == 100)

cat("Sweep grid:", nrow(sweep_grid), "valid ternary points.\n")

sweep_results <- classify_texture(sweep_grid, sand = sand, silt = silt,
                                  clay = clay)

na_count <- sum(is.na(sweep_results$.texture_class))
cat(sprintf("Points with NA class: %d / %d (%.1f%%)\n",
            na_count, nrow(sweep_grid), na_count / nrow(sweep_grid) * 100))

stopifnot("Unexpected NAs in ternary sweep" = na_count == 0)
cat("✓ All", nrow(sweep_grid), "points classified — including axis edges and vertices.\n")
```

```{r sweep-distribution}
sweep_results |>
  count(.texture_class) |>
  arrange(desc(n)) |>
  kable(caption = "Distribution of texture classes in the 5% ternary sweep")
```

### 1.7 Performance regression: 10 k points in < 1 s

The vectorised classification loop (iterates over 12 classes, tests all N
points per class simultaneously) must complete 10 000 samples in under one
second on any reasonable development machine.

```{r perf-regression}
set.seed(7743)
perf_data <- tibble::tibble(
  sand = runif(20000, 5, 80),
  clay = runif(20000, 5, 45)
) |>
  dplyr::mutate(silt = 100 - sand - clay) |>
  dplyr::filter(silt >= 5) |>
  dplyr::slice_head(n = 10000)

stopifnot("Perf grid is not 10 000 rows" = nrow(perf_data) == 10000)

elapsed <- system.time(
  perf_result <- classify_texture(perf_data,
                                  sand = sand, silt = silt, clay = clay)
)["elapsed"]

cat(sprintf("10 000 points classified in %.3f s\n", elapsed))
stopifnot("Performance regression: classify_texture() took > 1 s on 10 k points" =
            elapsed < 1)
cat("✓ Performance target met (< 1 s).\n")
```

```{r perf-na-check}
perf_na <- sum(is.na(perf_result$.texture_class))
cat(sprintf("NA classifications: %d / 10 000 (%.2f%%)\n",
            perf_na, perf_na / 100))
stopifnot("Unexpected NAs in performance grid" = perf_na == 0)
cat("✓ 0 NAs — all 10 000 points classified.\n")
```

---

## Part 2 — `gg_texture_triangle()`

### 2.1 Returns a ggplot object

```{r gg-returns-ggplot}
p <- gg_texture_triangle(
  class_centroids,
  sand = sand, silt = silt, clay = clay
)

stopifnot("Not a gg object" = inherits(p, "gg"))
cat("✓ gg_texture_triangle() returns a gg object.\n")
```

### 2.2 Colour mapping: NULL vs column name

```{r gg-colour-check}
# No colour mapping
p_no_colour <- gg_texture_triangle(class_centroids,
                                   sand = sand, silt = silt, clay = clay)
point_layer_nc <- Filter(\(l) inherits(l$geom, "GeomPoint"), p_no_colour$layers)
stopifnot("colour aesthetic present when it should be absent" =
            !"colour" %in% names(point_layer_nc[[1]]$mapping))
cat("✓ colour = NULL → no colour aesthetic on point layer.\n")

# With colour mapping
p_colour <- gg_texture_triangle(class_centroids,
                                sand = sand, silt = silt, clay = clay,
                                colour = expected_class)
point_layer_c <- Filter(\(l) inherits(l$geom, "GeomPoint"), p_colour$layers)
stopifnot("colour aesthetic absent when it should be present" =
            "colour" %in% names(point_layer_c[[1]]$mapping))
cat("✓ colour = expected_class → colour aesthetic present on point layer.\n")
```

### 2.3 Input validation

```{r gg-validation}
tryCatch(
  gg_texture_triangle(tibble(sand = 50, silt = 30, clay = 5),
                      sand = sand, silt = silt, clay = clay),
  error = \(e) cat("✓ Sum ≠ 100:", conditionMessage(e), "\n")
)

tryCatch(
  gg_texture_triangle(class_centroids,
                      sand = sand, silt = silt, clay = clay,
                      system = "BRITISH"),
  error = \(e) cat("✓ Bad system:", conditionMessage(e), "\n")
)
```

### 2.4 `breaks` parameter

Valid breaks produce grid lines; out-of-range values are rejected.

```{r breaks-valid}
# Quarters and deciles should both render without error
p_q <- gg_texture_triangle(class_centroids, sand, silt, clay,
                            breaks = c(25, 50, 75))
p_d <- gg_texture_triangle(class_centroids, sand, silt, clay,
                            breaks = seq(10, 90, 10))
stopifnot(inherits(p_q, "gg"), inherits(p_d, "gg"))
cat("✓ breaks = c(25,50,75) and seq(10,90,10) both produce gg objects.\n")
```

```{r breaks-invalid}
tryCatch(
  gg_texture_triangle(class_centroids, sand, silt, clay, breaks = c(0, 50)),
  error = \(e) cat("✓ breaks = 0 rejected:", conditionMessage(e), "\n")
)
tryCatch(
  gg_texture_triangle(class_centroids, sand, silt, clay, breaks = c(50, 100)),
  error = \(e) cat("✓ breaks = 100 rejected:", conditionMessage(e), "\n")
)
```

### 2.5 `style` parameter

`style = "none"` (default) leaves polygons unfilled; `style = "wiki"` maps hex
fills via `scale_fill_identity()`.

```{r style-layers}
p_none <- gg_texture_triangle(class_centroids, sand, silt, clay,
                               style = "none")
p_wiki <- gg_texture_triangle(class_centroids, sand, silt, clay,
                               style = "wiki")

# wiki plot should have a ScaleFilIdentity in its scales
has_fill_scale <- function(p) {
  any(vapply(p$scales$scales,
             \(s) inherits(s, "ScaleDiscreteIdentity") ||
                  inherits(s, "ScaleContinuousIdentity"),
             logical(1)))
}

stopifnot("wiki style missing fill scale" = has_fill_scale(p_wiki))
cat("✓ style = 'wiki' has a fill-identity scale; style = 'none' does not.\n")
```

```{r style-visual, fig.width=7, fig.height=6}
p_wiki + labs(title = "style = 'wiki'")
```

### 2.6 Grid styling (`grid_colour`, `grid_linetype`, `grid_linewidth`)

All combinations should produce gg objects without error.

```{r grid-styling}
combos <- list(
  list(col = "grey30", lt = "solid",  lw = 0.6),
  list(col = "grey80", lt = "dotted", lw = 0.3),
  list(col = "grey70", lt = "blank",  lw = 0.25)   # blank hides grid
)
for (co in combos) {
  p_g <- gg_texture_triangle(class_centroids, sand, silt, clay,
           grid_colour = co$col, grid_linetype = co$lt,
           grid_linewidth = co$lw)
  stopifnot(inherits(p_g, "gg"))
}
cat("✓ All grid_colour / grid_linetype / grid_linewidth combinations render.\n")
```

### 2.7 Border styling (`border_colour`, `border_linewidth`)

```{r border-styling}
# Bold visible borders
p_bold <- gg_texture_triangle(class_centroids, sand, silt, clay,
            style = "wiki", border_colour = "black", border_linewidth = 1.2)
stopifnot(inherits(p_bold, "gg"))

# NA border (hidden) should not error
p_noborder <- gg_texture_triangle(class_centroids, sand, silt, clay,
               style = "wiki", border_colour = NA)
stopifnot(inherits(p_noborder, "gg"))

cat("✓ border_colour (including NA) and border_linewidth render without error.\n")
```

### 2.8 Tick height (`tick_height`)

```{r tick-height}
# Zero hides tick marks
p_t0 <- gg_texture_triangle(class_centroids, sand, silt, clay, tick_height = 0)
# Fixed height
p_tf <- gg_texture_triangle(class_centroids, sand, silt, clay, tick_height = 0.05)
# Negative should error
tryCatch(
  gg_texture_triangle(class_centroids, sand, silt, clay, tick_height = -0.01),
  error = \(e) cat("✓ Negative tick_height rejected:", conditionMessage(e), "\n")
)
stopifnot(inherits(p_t0, "gg"), inherits(p_tf, "gg"))
cat("✓ tick_height = 0 and fixed height both render.\n")
```

### 2.9 `texture_label_style()` — validation and key filtering

```{r label-style-valid}
# Only supplied keys should appear in the output list
ls1 <- texture_label_style(class_show = FALSE, axis_size = 4.5)
stopifnot(
  "class_show not in output" = "class_show" %in% names(ls1),
  "axis_size not in output"  = "axis_size"  %in% names(ls1),
  "unsupplied key leaked"    = !"tick_show" %in% names(ls1),
  "class_show value wrong"   = ls1$class_show == FALSE,
  "axis_size value wrong"    = ls1$axis_size  == 4.5
)
cat("✓ texture_label_style() returns only explicitly supplied keys.\n")
```

```{r label-style-errors}
tryCatch(
  texture_label_style(tick_show = "yes"),
  error = \(e) cat("✓ Non-logical tick_show rejected:", conditionMessage(e), "\n")
)
tryCatch(
  texture_label_style(axis_size = -1),
  error = \(e) cat("✓ Negative axis_size rejected:", conditionMessage(e), "\n")
)
tryCatch(
  texture_label_style(class_size = 0),
  error = \(e) cat("✓ Zero class_size rejected:", conditionMessage(e), "\n")
)
```

```{r label-style-integration}
# All six keys in a single call, passed via label_style — should render
p_ls <- gg_texture_triangle(class_centroids, sand, silt, clay,
  label_style = texture_label_style(
    tick_show  = FALSE, tick_size  = 2.0,
    axis_show  = TRUE,  axis_size  = 4.0,
    class_show = TRUE,  class_size = 3.0
  ))
stopifnot(inherits(p_ls, "gg"))
cat("✓ All six label_style keys accepted and rendered.\n")
```

### 2.10 `texture_surface()` — constructor validation

```{r surface-setup}
set.seed(5512)
surf_df <- tibble::tibble(
  sand = runif(70, 5, 85),
  clay = runif(70, 5, 50)
) |>
  dplyr::mutate(silt = 100 - sand - clay) |>
  dplyr::filter(silt >= 2) |>
  dplyr::mutate(prop = clay * 0.6 + rnorm(dplyr::n(), 0, 3),
                label = "x")
```

```{r surface-valid}
ts_obj <- texture_surface(surf_df, sand, silt, clay, z = prop)
stopifnot(
  "wrong class"          = inherits(ts_obj, "texture_surface"),
  "missing sand_v"       = "sand_v"     %in% names(ts_obj),
  "missing z_v"          = "z_v"        %in% names(ts_obj),
  "wrong resolution"     = ts_obj$resolution == 150L,
  "wrong power"          = ts_obj$power == 2,
  "z length mismatch"    = length(ts_obj$z_v) == length(ts_obj$sand_v)
)
cat("✓ texture_surface() returns a valid 'texture_surface' object.\n")
```

```{r surface-errors}
# Non-numeric z
tryCatch(
  texture_surface(surf_df, sand, silt, clay, z = label),
  error = \(e) cat("✓ Non-numeric z rejected:", conditionMessage(e), "\n")
)
# Resolution too small
tryCatch(
  texture_surface(surf_df, sand, silt, clay, z = prop, resolution = 5L),
  error = \(e) cat("✓ resolution < 10 rejected:", conditionMessage(e), "\n")
)
# Bad power
tryCatch(
  texture_surface(surf_df, sand, silt, clay, z = prop, power = -1),
  error = \(e) cat("✓ Negative power rejected:", conditionMessage(e), "\n")
)
# Wrong class passed as surface
tryCatch(
  gg_texture_triangle(class_centroids, sand, silt, clay,
                      surface = list(a = 1)),
  error = \(e) cat("✓ Non-texture_surface object rejected:", conditionMessage(e), "\n")
)
```

```{r surface-renders}
p_surf <- gg_texture_triangle(class_centroids, sand, silt, clay,
            surface = texture_surface(surf_df, sand, silt, clay, z = prop)) +
          ggplot2::scale_fill_viridis_c(name = "prop")
stopifnot(inherits(p_surf, "gg"))
cat("✓ gg_texture_triangle() with surface renders a gg object.\n")
```

### 2.11 `geom_texture_contour()` — layer output

```{r contour-layer}
contour_layer <- geom_texture_contour(surf_df, sand, silt, clay, z = prop,
                                      breaks = 20)
stopifnot("Not a ggplot2 Layer" = inherits(contour_layer, "LayerInstance"))
cat("✓ geom_texture_contour() returns a ggplot2 LayerInstance.\n")
```

```{r contour-renders}
# Addable to a plain triangle
p_ct <- gg_texture_triangle(class_centroids, sand, silt, clay) +
        geom_texture_contour(surf_df, sand, silt, clay, z = prop,
                             breaks = c(15, 25), colour = "steelblue",
                             linewidth = 0.9)
stopifnot(inherits(p_ct, "gg"))

# Addable on top of a surface
p_cs <- gg_texture_triangle(class_centroids, sand, silt, clay,
          surface = texture_surface(surf_df, sand, silt, clay, z = prop)) +
        ggplot2::scale_fill_viridis_c() +
        geom_texture_contour(surf_df, sand, silt, clay, z = prop,
                             breaks = 20, colour = "white", linewidth = 1.2)
stopifnot(inherits(p_cs, "gg"))
cat("✓ geom_texture_contour() composable on plain triangle and surface.\n")
```

---

### 2.12 Expand-grid visualisation: all 12 class centroids, coloured by class

```{r gg-centroids, fig.width=7, fig.height=6.5}
gg_texture_triangle(
  class_centroids,
  sand   = sand,
  silt   = silt,
  clay   = clay,
  colour = expected_class
) +
  labs(
    title   = "USDA texture triangle — 12 class centroids",
    colour  = "Expected class"
  )
```

### 2.13 Expand-grid visualisation: 60-sample grid, coloured by correct/incorrect

```{r gg-multisamples, fig.width=7, fig.height=6.5}
gg_texture_triangle(
  sample_results |>
    mutate(result = if_else(
      expected_class == .texture_class, "correct", "incorrect"
    )),
  sand   = sand,
  silt   = silt,
  clay   = clay,
  colour = result
) +
  scale_colour_manual(values = c(correct = "steelblue", incorrect = "tomato")) +
  labs(
    title   = "60-sample grid — classification correctness",
    colour  = "Result"
  )
```

### 2.14 Expand-grid visualisation: systematic 5 % sweep coloured by class

```{r gg-sweep, fig.width=7, fig.height=6.5}
gg_texture_triangle(
  sweep_results,
  sand   = sand,
  silt   = silt,
  clay   = clay,
  colour = .texture_class
) +
  labs(
    title   = "Systematic ternary sweep (5 % grid) — texture classification",
    colour  = "Texture class"
  )
```

---

## Part 3 — SpatRaster large-raster benchmark

Simulates a 1 000 × 1 000 SpatRaster (1 M cells) with three layers (sand,
silt, clay) drawn from realistic ternary fractions. Benchmarks
`classify_texture.SpatRaster()` end-to-end and confirms 0 NA classifications
in valid cells.

```{r spatraster-bench-setup}
if (!requireNamespace("terra", quietly = TRUE)) {
  message("terra not installed — skipping Part 3 benchmark.")
  knitr::knit_exit()
}
library(terra)
```

### 3.1 Build 1 M-cell SpatRaster

```{r build-1m-raster}
set.seed(4421)
n_cells <- 1000L * 1000L   # 1 M cells

# Generate valid ternary fractions: sand + silt + clay = 100
sand_v <- runif(n_cells, 5, 80)
clay_v <- runif(n_cells, 5, 45)
silt_v <- 100 - sand_v - clay_v

# Clamp so silt >= 2 (should already hold for most draws; discard tiny fraction)
valid <- silt_v >= 2
cat(sprintf("Valid cells: %d / %d (%.2f%%)\n",
            sum(valid), n_cells, mean(valid) * 100))

# Keep first 1 M valid (almost certainly all of them; just trim if needed)
sand_v <- sand_v[valid][seq_len(min(n_cells, sum(valid)))]
silt_v <- silt_v[valid][seq_len(min(n_cells, sum(valid)))]
clay_v <- clay_v[valid][seq_len(min(n_cells, sum(valid)))]
n_actual <- length(sand_v)

# Build a minimal 1000×1000 raster with CRS
r_bench <- terra::rast(ncols = 1000, nrows = 1000, nlyrs = 3,
                       xmin = 0, xmax = 1, ymin = 0, ymax = 1,
                       crs = "EPSG:4326")
names(r_bench) <- c("sand", "silt", "clay")
terra::values(r_bench) <- cbind(sand_v, silt_v, clay_v)

cat(sprintf("Raster: %d cols × %d rows × 3 layers (%d cells total)\n",
            terra::ncol(r_bench), terra::nrow(r_bench), terra::ncell(r_bench)))
```

### 3.2 Time `classify_texture()` on 1 M cells

```{r bench-1m-classify}
library(tidysoiltexture)

elapsed_1m <- system.time(
  r_class <- classify_texture(r_bench,
                              sand = "sand", silt = "silt", clay = "clay")
)["elapsed"]

cat(sprintf("1 M cells classified in %.3f s\n", elapsed_1m))
stopifnot("Performance regression: classify_texture() took > 15 s on 1 M cells" =
            elapsed_1m < 15)
cat("✓ Performance target met (< 15 s).\n")
```

### 3.3 Verify output — 0 NAs in valid cells

```{r bench-na-check}
out_vals  <- terra::values(r_class)
n_na      <- sum(is.na(out_vals))
pct_na    <- n_na / n_actual * 100
cat(sprintf("NA cells: %d / %d (%.4f%%)\n", n_na, n_actual, pct_na))
stopifnot("Unexpected NAs in 1 M-cell benchmark" = n_na == 0L)
cat("✓ 0 NAs — all 1 M cells classified.\n")
```

### 3.4 Class distribution

```{r bench-distribution}
class_levels <- terra::levels(r_class)[[1]]
tbl_int      <- as.vector(terra::values(r_class))

tibble::tibble(id = tbl_int) |>
  dplyr::left_join(class_levels, by = "id") |>
  dplyr::count(texture_class) |>
  dplyr::mutate(pct = round(n / sum(n) * 100, 2)) |>
  dplyr::arrange(desc(n)) |>
  knitr::kable(caption = "Class distribution across 1 M-cell raster")
```

### 3.5 Visual: classified raster

```{r bench-plot, fig.width=6, fig.height=5}
terra::plot(r_class,
            main   = "Texture classification — 1 M-cell benchmark raster",
            col    = grDevices::hcl.colors(12, "Set2"),
            legend = TRUE)
```

---

## Summary

```{r summary}
tibble::tribble(
  ~Function,                   ~`Grid / input`,                        ~`Tests run`,
  "classify_texture()",        "12 centroids",
    "correct class & abbr, output structure",
  "classify_texture()",        "60-sample grid (5 per class)",
    "confusion matrix, per-class accuracy, NA coverage",
  "classify_texture()",        "ternary sweep at 5 % intervals (231 pts)",
    "0 NAs incl. axis edges & vertices, class distribution",
  "classify_texture()",        "equivalence & validation",
    "column vs scalar, sum ≠ 100, unsupported system",
  "classify_texture()",        "10 k-point performance regression",
    "elapsed < 1 s; 0 NAs",
  "gg_texture_triangle()",     "gg object & colour aesthetic",
    "inherits gg, colour mapped/absent correctly",
  "gg_texture_triangle()",     "breaks parameter",
    "valid combos render; 0 and 100 rejected",
  "gg_texture_triangle()",     "style parameter",
    "wiki fill scale present; none absent; visual render",
  "gg_texture_triangle()",     "grid_colour/linetype/linewidth",
    "all combinations render without error",
  "gg_texture_triangle()",     "border_colour/border_linewidth",
    "bold & NA borders both accepted",
  "gg_texture_triangle()",     "tick_height",
    "0 (hidden) and fixed height render; negative rejected",
  "texture_label_style()",     "key filtering & validation",
    "only set keys returned; bad types/values rejected; all 6 keys work",
  "texture_surface()",         "constructor & validation",
    "correct class/structure; non-numeric z, bad resolution/power rejected",
  "geom_texture_contour()",    "layer output & composability",
    "returns LayerInstance; addable to plain triangle and surface",
  "gg_texture_triangle()",     "12 centroids, 60-sample, sweep",
    "visualisations — colour by class, correctness, texture class",
  "classify_texture.SpatRaster()", "1 M-cell SpatRaster (1000×1000)",
    "0 NAs; elapsed < 15 s; class distribution table; visual plot"
) |>
  kable(caption = "Test coverage summary")
```
