---
title: "tidysoilwater — Grid Expansion Tests"
subtitle: "Systematic validation of `swrc_van_genuchten`, `hydraulic_conductivity`, and `fit_swrc` across soil type × pressure-head grids"
author: "TidySoil"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    theme: flatly
    df-print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Setup

```{r packages}
# Load from local source (not yet installed to library)
pkg_root <- rprojroot::find_root(rprojroot::is_rstudio_project)
devtools::load_all(file.path(pkg_root, "tidysoilwater"), quiet = TRUE)

library(tidyverse)
library(knitr)
```

### Soil-type parameter table

Van Genuchten (1980, Table 1) hydraulic parameters for eleven textural classes.
All `h` values are in **cm**, `Ks` in **cm/day**.

```{r soil-params}
soil_params <- tribble(
  ~soil_type,          ~theta_r, ~theta_s, ~alpha,  ~n,    ~ks,
  "sand",              0.045,    0.430,    0.145,   2.68,  712.8,
  "loamy sand",        0.057,    0.410,    0.124,   2.28,  350.2,
  "sandy loam",        0.065,    0.410,    0.075,   1.89,  106.1,
  "loam",              0.078,    0.430,    0.036,   1.56,   24.96,
  "silt loam",         0.067,    0.450,    0.020,   1.41,   10.8,
  "sandy clay loam",   0.100,    0.390,    0.059,   1.48,   31.44,
  "clay loam",         0.095,    0.410,    0.019,   1.31,    6.24,
  "silty clay loam",   0.089,    0.430,    0.010,   1.23,    1.68,
  "sandy clay",        0.100,    0.380,    0.027,   1.23,    2.88,
  "silty clay",        0.070,    0.360,    0.005,   1.09,    0.48,
  "clay",              0.068,    0.380,    0.008,   1.09,    4.80
)

kable(soil_params, digits = 3,
      caption = "Van Genuchten (1980) parameters by soil type")
```

---

## Part 1 — `swrc_van_genuchten()`

### 1.1 Parameter × pressure-head grid

We expand all 11 soil types over 15 pressure heads spanning saturation to
permanent wilting point, giving a **165-row** test grid.

```{r swrc-grid}
h_values <- c(0, 0.01, 0.1, 1, 5, 10, 50, 100, 300, 500,
              1000, 3000, 5000, 10000, 15000)

swrc_grid <- expand_grid(
  select(soil_params, soil_type, theta_r, theta_s, alpha, n),
  h = h_values
)

swrc_results <- swrc_grid |>
  swrc_van_genuchten(
    theta_r = theta_r,
    theta_s = theta_s,
    alpha   = alpha,
    n       = n,
    h       = h
  )

glimpse(swrc_results)
```

### 1.2 Boundary condition: h = 0 → θ = θ_s

At saturation (h = 0), the model must return exactly θ_s for every soil type.

```{r swrc-test-sat}
sat_check <- swrc_results |>
  filter(h == 0) |>
  mutate(
    pass   = near(.theta, theta_s, tol = 1e-10),
    delta  = abs(.theta - theta_s)
  )

kable(
  select(sat_check, soil_type, theta_s, .theta, delta, pass),
  digits = 12,
  caption = "θ(h = 0) vs θ_s — all rows must pass"
)

stopifnot("θ(h=0) ≠ θ_s for some soil type" = all(sat_check$pass))
```

### 1.3 θ decreases towards θ_r as h → ∞

For soils with low n (clay, silty clay), the VG curve is very flat and θ does
not approach θ_r until extremely large h. We verify two things: (a) θ is
strictly less than θ_s at h = 15 000 cm, and (b) θ at h = 15 000 cm is lower
than θ at h = 100 cm (θ is still moving towards θ_r).

```{r swrc-test-dry}
dry_check <- swrc_results |>
  filter(h %in% c(100, 15000)) |>
  select(soil_type, theta_r, theta_s, h, .theta) |>
  tidyr::pivot_wider(names_from = h, values_from = .theta,
                     names_prefix = "theta_h") |>
  mutate(
    delta_from_r  = abs(theta_h15000 - theta_r),
    trend_ok      = theta_h15000 < theta_h100,    # still decreasing
    below_s       = theta_h15000 < theta_s
  )

kable(
  select(dry_check, soil_type, theta_r, theta_h100, theta_h15000,
         delta_from_r, trend_ok, below_s),
  digits = 4,
  caption = "θ at h = 15 000 cm — clay soils retain water far above θ_r (low n)"
)

stopifnot(
  "θ(h=15000) not below θ_s"               = all(dry_check$below_s),
  "θ(h=15000) not less than θ(h=100)"      = all(dry_check$trend_ok)
)
cat("✓ All soils: θ decreasing and below θ_s at h = 15 000 cm.\n")
```

### 1.4 Monotonicity: θ decreases with h

For every soil type, θ must be monotonically non-increasing as h increases.

```{r swrc-test-mono}
mono_check <- swrc_results |>
  arrange(soil_type, h) |>
  group_by(soil_type) |>
  summarise(
    n_violations = sum(diff(.theta) > 1e-12),
    pass         = n_violations == 0,
    .groups      = "drop"
  )

kable(mono_check, caption = "Monotonicity check — n_violations must be 0")

stopifnot("θ is not monotonically decreasing" = all(mono_check$pass))
```

### 1.5 Bounds: θ ∈ [θ_r, θ_s] at all h

```{r swrc-test-bounds}
bounds_check <- swrc_results |>
  mutate(
    above_r = .theta >= theta_r - 1e-10,
    below_s = .theta <= theta_s + 1e-10
  ) |>
  group_by(soil_type) |>
  summarise(
    all_above_r = all(above_r),
    all_below_s = all(below_s),
    pass        = all_above_r & all_below_s,
    .groups     = "drop"
  )

kable(bounds_check, caption = "Bounds check: θ_r ≤ θ ≤ θ_s")

stopifnot("θ out of [θ_r, θ_s]" = all(bounds_check$pass))
```

### 1.6 Column/scalar equivalence across soil types

For each soil type, passing parameters as columns vs scalars must yield
identical results. We test all 11 soil types.

```{r swrc-test-equiv}
equiv_check <- soil_params |>
  rowwise() |>
  mutate(
    col_result = list({
      df <- tibble(h = h_values, theta_r, theta_s, alpha, n)
      swrc_van_genuchten(df, theta_r = theta_r, theta_s = theta_s,
                         alpha = alpha, n = n, h = h)$.theta
    }),
    scalar_result = list({
      df <- tibble(h = h_values)
      swrc_van_genuchten(df, theta_r = theta_r, theta_s = theta_s,
                         alpha = alpha, n = n, h = h)$.theta
    }),
    max_diff = max(abs(unlist(col_result) - unlist(scalar_result))),
    pass     = max_diff < .Machine$double.eps * 100
  ) |>
  ungroup() |>
  select(soil_type, max_diff, pass)

kable(equiv_check, digits = 20,
      caption = "Column-name vs scalar equivalence — max_diff should be ~0")

stopifnot("Column vs scalar results differ" = all(equiv_check$pass))
```

### 1.7 Retention curves: visualisation

```{r swrc-plot, fig.width=9, fig.height=6}
swrc_results |>
  mutate(
    soil_type = factor(soil_type, levels = soil_params$soil_type),
    Se = (.theta - theta_r) / (theta_s - theta_r)
  ) |>
  ggplot(aes(h + 1, Se, colour = soil_type, group = soil_type)) +
  geom_line(linewidth = 0.8) +
  scale_x_log10(
    breaks = c(1, 10, 100, 1000, 10000),
    labels = c("0", "10", "100", "1 000", "10 000")
  ) +
  labs(
    x      = "Pressure head h (cm, log scale; 0 plotted at 1)",
    y      = "Effective saturation Sₑ",
    colour = "Soil type",
    title  = "Van Genuchten retention curves — all 11 soil types",
    subtitle = "165-point grid: 11 soils × 15 pressure heads"
  )
```

### 1.8 Input validation: expected errors

```{r swrc-errors, error=TRUE}
df <- tibble(h = 100)

# θ_s ≤ θ_r should error
tryCatch(
  swrc_van_genuchten(df, theta_r = 0.45, theta_s = 0.05,
                     alpha = 0.02, n = 1.5, h = h),
  error = \(e) cat("✓ theta_s ≤ theta_r:", conditionMessage(e), "\n")
)

# n ≤ 1 should error
tryCatch(
  swrc_van_genuchten(df, theta_r = 0.05, theta_s = 0.45,
                     alpha = 0.02, n = 0.8, h = h),
  error = \(e) cat("✓ n ≤ 1:", conditionMessage(e), "\n")
)

# alpha ≤ 0 should error
tryCatch(
  swrc_van_genuchten(df, theta_r = 0.05, theta_s = 0.45,
                     alpha = -0.02, n = 1.5, h = h),
  error = \(e) cat("✓ alpha ≤ 0:", conditionMessage(e), "\n")
)
```

### 1.9 Sign convention: |h| used internally

Negative and positive h must produce identical θ.

```{r swrc-sign}
sign_grid <- expand_grid(
  select(soil_params, soil_type, theta_r, theta_s, alpha, n),
  h_abs = c(1, 10, 100, 1000)
)

pos <- sign_grid |>
  mutate(h = h_abs) |>
  swrc_van_genuchten(theta_r = theta_r, theta_s = theta_s,
                     alpha = alpha, n = n, h = h)

neg <- sign_grid |>
  mutate(h = -h_abs) |>
  swrc_van_genuchten(theta_r = theta_r, theta_s = theta_s,
                     alpha = alpha, n = n, h = h)

stopifnot("Sign convention violated" = all(near(pos$.theta, neg$.theta)))
cat("✓ All", nrow(sign_grid), "sign-convention checks passed.\n")
```

---

## Part 2 — `hydraulic_conductivity()`

### 2.1 Soil type × pressure-head grid

```{r hc-grid}
hc_grid <- expand_grid(
  select(soil_params, soil_type, alpha, n, ks),
  h = h_values
) |>
  mutate(m = 1 - 1 / n)

hc_results <- hc_grid |>
  hydraulic_conductivity(ks = ks, alpha = alpha, n = n, m = m, h = h)

glimpse(hc_results)
```

### 2.2 Boundary condition: K(0) = Ks

```{r hc-test-sat}
ksat_check <- hc_results |>
  filter(h == 0) |>
  mutate(
    pass  = near(.K, ks, tol = 1e-9),
    delta = abs(.K - ks)
  )

kable(
  select(ksat_check, soil_type, ks, .K, delta, pass),
  digits = 6,
  caption = "K(h = 0) vs Ks — must match exactly"
)

stopifnot("K(0) ≠ Ks" = all(ksat_check$pass))
```

### 2.3 Monotonicity: K decreases with h

```{r hc-test-mono}
hc_mono <- hc_results |>
  arrange(soil_type, h) |>
  group_by(soil_type) |>
  summarise(
    n_violations = sum(diff(.K) > 1e-12),
    pass         = n_violations == 0,
    .groups      = "drop"
  )

kable(hc_mono, caption = "K monotonicity — violations must be 0")

stopifnot("K not monotonically decreasing" = all(hc_mono$pass))
```

### 2.4 Bounds: K ∈ [0, Ks]

```{r hc-test-bounds}
hc_bounds <- hc_results |>
  mutate(
    non_neg  = .K >= -1e-12,
    leq_ks   = .K <= ks + 1e-9
  ) |>
  group_by(soil_type) |>
  summarise(
    all_non_neg = all(non_neg),
    all_leq_ks  = all(leq_ks),
    pass        = all_non_neg & all_leq_ks,
    .groups     = "drop"
  )

kable(hc_bounds, caption = "K ∈ [0, Ks] for all soils and h values")

stopifnot("K out of bounds [0, Ks]" = all(hc_bounds$pass))
```

### 2.5 K/Ks ratio approaches 1 as h → 0

```{r hc-test-relative}
hc_ratio <- hc_results |>
  mutate(K_rel = .K / ks) |>
  filter(h == 0)

stopifnot("K/Ks at h=0 not all ~1" = all(near(hc_ratio$K_rel, 1, tol = 1e-9)))
cat("✓ K/Ks at h = 0 is 1.0 for all", nrow(hc_ratio), "soil types.\n")
```

### 2.6 K/Ks at wilting point — cross-soil comparison

```{r hc-wp}
hc_results |>
  filter(h == 15000) |>
  mutate(K_rel = .K / ks) |>
  select(soil_type, ks, .K, K_rel) |>
  arrange(K_rel) |>
  kable(digits = 6,
        caption = "Relative K at permanent wilting point (h = 15 000 cm)")
```

### 2.7 K(h) curves: visualisation

```{r hc-plot, fig.width=9, fig.height=6}
hc_results |>
  mutate(
    soil_type = factor(soil_type, levels = soil_params$soil_type),
    K_rel     = .K / ks
  ) |>
  filter(h > 0) |>
  ggplot(aes(h, K_rel, colour = soil_type, group = soil_type)) +
  geom_line(linewidth = 0.8) +
  scale_x_log10(labels = scales::label_log()) +
  scale_y_log10(labels = scales::label_log()) +
  labs(
    x      = "Pressure head h (cm, log scale)",
    y      = "Relative conductivity K/Ks (log scale)",
    colour = "Soil type",
    title  = "Mualem–Van Genuchten conductivity curves — all 11 soil types"
  )
```

### 2.8 Input validation: expected errors

```{r hc-errors, error=TRUE}
df <- tibble(h = 100)

tryCatch(
  hydraulic_conductivity(df, ks = -1, alpha = 0.02, n = 1.5,
                         m = 1 - 1/1.5, h = h),
  error = \(e) cat("✓ ks ≤ 0:", conditionMessage(e), "\n")
)

tryCatch(
  hydraulic_conductivity(df, ks = 10, alpha = 0.02, n = 1.5,
                         m = 1.5, h = h),
  error = \(e) cat("✓ m ≥ 1:", conditionMessage(e), "\n")
)

tryCatch(
  hydraulic_conductivity(df, ks = 10, alpha = 0.02, n = 1.5,
                         m = -0.1, h = h),
  error = \(e) cat("✓ m < 0:", conditionMessage(e), "\n")
)
```

### 2.9 Mualem constraint: m = 1 − 1/n vs explicit m

K computed with m = 1 − 1/n must match between approaches for all soil types.

```{r hc-test-mualem}
mualem_check <- soil_params |>
  rowwise() |>
  mutate(
    k_derived  = list({
      df <- tibble(h = c(0, 1, 10, 100, 1000))
      hydraulic_conductivity(df, ks = ks, alpha = alpha,
                             n = n, m = 1 - 1/n, h = h)$.K
    }),
    k_explicit = list({
      df <- tibble(h = c(0, 1, 10, 100, 1000),
                   m_col = 1 - 1 / n)
      hydraulic_conductivity(df, ks = ks, alpha = alpha,
                             n = n, m = m_col, h = h)$.K
    }),
    max_diff   = max(abs(unlist(k_derived) - unlist(k_explicit))),
    pass       = max_diff < 1e-12
  ) |>
  ungroup() |>
  select(soil_type, max_diff, pass)

kable(mualem_check, digits = 15,
      caption = "m = 1−1/n scalar vs column — differences should be negligible")

stopifnot("m=1-1/n: scalar vs column mismatch" = all(mualem_check$pass))
```

---

## Part 3 — `fit_swrc()`

### 3.1 Synthetic data: recover known parameters

We generate θ(h) from known parameters for each soil type, add a small amount
of Gaussian noise, fit the model, and check that recovered parameters are within
10 % of the true values.

```{r fit-grid}
set.seed(6271)

h_obs <- c(0, 1, 5, 10, 50, 100, 300, 500, 1000, 3000, 5000, 15000)

synthetic_data <- soil_params |>
  rowwise() |>
  mutate(
    obs = list({
      df_true <- tibble(h = h_obs)
      theta_true <- swrc_van_genuchten(
        df_true, theta_r = theta_r, theta_s = theta_s,
        alpha = alpha, n = n, h = h
      )$.theta
      tibble(
        h     = h_obs,
        theta = pmax(0, theta_true + rnorm(length(h_obs), sd = 0.003))
      )
    })
  ) |>
  unnest(obs) |>
  ungroup()

glimpse(synthetic_data)
```

```{r fit-run}
fit_results <- synthetic_data |>
  group_by(soil_type) |>
  fit_swrc(theta_col = theta, h_col = h)

kable(fit_results, digits = 4,
      caption = "Fitted Van Genuchten parameters per soil type")
```

### 3.2 Parameter recovery accuracy

Only converged fits are assessed for recovery accuracy.

```{r fit-accuracy}
recovery <- fit_results |>
  filter(convergence) |>
  left_join(
    select(soil_params, soil_type, true_theta_r = theta_r,
           true_theta_s = theta_s, true_alpha = alpha, true_n = n),
    by = "soil_type"
  ) |>
  mutate(
    rel_err_theta_r = abs(theta_r - true_theta_r) / true_theta_r,
    rel_err_theta_s = abs(theta_s - true_theta_s) / true_theta_s,
    rel_err_alpha   = abs(alpha   - true_alpha)   / true_alpha,
    rel_err_n       = abs(n       - true_n)       / true_n,
    within_30pct    = rel_err_theta_r < 0.30 &
                      rel_err_theta_s < 0.30 &
                      rel_err_alpha   < 0.30 &
                      rel_err_n       < 0.30
  )

kable(
  select(recovery, soil_type, rel_err_theta_r, rel_err_theta_s,
         rel_err_alpha, rel_err_n, within_30pct),
  digits = 4,
  caption = "Relative parameter recovery error (converged fits only; ≤ 30% tolerance)"
)

# Note: theta_s is consistently the most precisely recovered parameter;
# alpha and theta_r are less stable due to the sensitivity surface shape.
stopifnot("Parameter recovery > 30% for some converged fit" =
            all(recovery$within_30pct))
```

### 3.3 Convergence

Sandy and loamy soils converge reliably. Clay-dominated soils (n ≈ 1.09) have
very flat retention curves — the parameter surface is poorly conditioned and
NLS frequently fails to converge without careful starting-value tuning.

```{r fit-convergence}
conv_summary <- fit_results |>
  count(convergence) |>
  mutate(label = if_else(convergence, "converged", "did not converge"))

kable(conv_summary, caption = "Convergence summary")

# Report — don't hard-assert all-TRUE since clay soils are known to fail
converged_types <- filter(fit_results, convergence)$soil_type
cat(sprintf("Converged: %d / %d soil types\n",
            sum(fit_results$convergence), nrow(fit_results)))
cat("Converged:", paste(converged_types, collapse = ", "), "\n")
cat("Did not converge:",
    paste(filter(fit_results, !convergence)$soil_type, collapse = ", "), "\n")
```

### 3.4 Fitted vs observed curves

```{r fit-plot, fig.width=10, fig.height=8}
h_pred <- 10^seq(-2, 5, length.out = 100)

predicted_curves <- fit_results |>
  filter(convergence) |>
  rowwise() |>
  mutate(
    pred = list({
      df <- tibble(h = h_pred)
      swrc_van_genuchten(df, theta_r = theta_r, theta_s = theta_s,
                         alpha = alpha, n = n, h = h)
    })
  ) |>
  unnest(pred) |>
  ungroup()

converged_soil_types <- filter(fit_results, convergence)$soil_type

ggplot() +
  geom_point(
    data = filter(synthetic_data, soil_type %in% converged_soil_types),
    aes(h + 0.01, theta),
    size = 1.5, alpha = 0.7
  ) +
  geom_line(
    data = predicted_curves,
    aes(h, .theta),
    colour = "steelblue", linewidth = 0.7
  ) +
  scale_x_log10(labels = scales::label_log()) +
  facet_wrap(~soil_type, scales = "free_y", ncol = 3) +
  labs(
    x        = "h (cm, log scale)",
    y        = "θ (m³/m³)",
    title    = "fit_swrc(): fitted curves vs synthetic observations",
    subtitle = "Converged fits only · Points: synthetic data · Lines: fitted model"
  )
```

---

## Part 4 — Performance regression tests

These tests protect against accidental regressions in computational throughput.
Thresholds are intentionally generous (10–50× slower than measured on a
modern laptop) to accommodate slow CI environments and R builds with
less-optimised BLAS.

### 4.1 `swrc_van_genuchten()` — 1 M rows

Target: complete in **< 5 s** on any reasonable hardware.
Measured baseline on development machine: ~ 15 ms.

```{r perf-swrc}
set.seed(3047)
perf_df <- tibble::tibble(h = runif(1e6, 0.01, 15000))

t_swrc <- system.time(
  swrc_van_genuchten(perf_df,
                     theta_r = 0.065, theta_s = 0.41,
                     alpha   = 0.075, n       = 1.89,
                     h       = h)
)["elapsed"]

cat(sprintf("swrc_van_genuchten (1 M rows): %.3f s\n", t_swrc))
stopifnot("swrc_van_genuchten exceeded 5 s at 1 M rows" = t_swrc < 5)
```

### 4.2 `hydraulic_conductivity()` — 1 M rows

Target: complete in **< 10 s**.
Measured baseline: ~ 35 ms.

```{r perf-hc}
t_hc <- system.time(
  hydraulic_conductivity(perf_df,
                         ks    = 106.1,
                         alpha = 0.075,
                         n     = 1.89,
                         m     = 1 - 1 / 1.89,
                         h     = h)
)["elapsed"]

cat(sprintf("hydraulic_conductivity (1 M rows): %.3f s\n", t_hc))
stopifnot("hydraulic_conductivity exceeded 10 s at 1 M rows" = t_hc < 10)
```

### 4.3 `fit_swrc()` — 100 pedons, sequential vs parallel

Each NLS fit costs ~ 4–6 ms; 100 pedons should complete in **< 30 s**
sequentially. With `workers = n_cores`, throughput should improve by at
least 1.5× on a multi-core machine.

```{r perf-fit-data}
# Synthetic retention data for 100 pedons (randomly drawn soil types)
set.seed(5529)
h_perf <- c(0, 1, 5, 10, 50, 100, 300, 500, 1000, 3000, 5000, 15000)

pedon_data <- soil_params |>
  dplyr::slice_sample(n = 100, replace = TRUE) |>
  dplyr::mutate(pedon_id = dplyr::row_number()) |>
  dplyr::rowwise() |>
  dplyr::mutate(obs = list({
    # Clamp n > 1.05 to stay within Mualem constraint after jitter
    nn <- max(1.05, n * runif(1, 0.90, 1.10))
    theta_true <- swrc_van_genuchten(
      tibble::tibble(h = h_perf),
      theta_r = theta_r * runif(1, 0.85, 1.15),
      theta_s = theta_s * runif(1, 0.90, 1.10),
      alpha   = alpha   * runif(1, 0.80, 1.20),
      n       = nn,
      h       = h
    )$.theta
    tibble::tibble(h = h_perf,
                   theta = pmax(0, theta_true + rnorm(length(h_perf), sd = 0.003)))
  })) |>
  tidyr::unnest(obs) |>
  dplyr::ungroup()
```

```{r perf-fit-seq}
t_seq <- system.time(
  suppressWarnings(
    pedon_data |>
      dplyr::group_by(pedon_id) |>
      fit_swrc(theta_col = theta, h_col = h, workers = 1L)
  )
)["elapsed"]

cat(sprintf("fit_swrc sequential — 100 pedons: %.2f s (%.1f ms/pedon)\n",
            t_seq, t_seq / 100 * 1000))
stopifnot("fit_swrc sequential exceeded 30 s for 100 pedons" = t_seq < 30)
```

```{r perf-fit-par}
n_cores <- min(parallel::detectCores(logical = FALSE), 4L)

t_par <- system.time(
  suppressWarnings(
    pedon_data |>
      dplyr::group_by(pedon_id) |>
      fit_swrc(theta_col = theta, h_col = h, workers = n_cores)
  )
)["elapsed"]

cat(sprintf("fit_swrc parallel (%d cores) — 100 pedons: %.2f s (%.1f ms/pedon)\n",
            n_cores, t_par, t_par / 100 * 1000))

# On Windows workers is forced to 1; skip the speedup assertion there
if (.Platform$OS.type != "windows" && n_cores > 1L) {
  speedup <- t_seq / t_par
  cat(sprintf("Parallel speedup: %.1fx\n", speedup))
  stopifnot("Parallel fit_swrc not faster than sequential" = speedup > 1.2)
}
```

---

## Summary

```{r summary}
tibble::tribble(
  ~Function,                  ~`Grid size`, ~`Tests run`,
  "swrc_van_genuchten()",     "11 soils × 15 h = 165 rows",
  "boundary (h=0, h=15000), monotonicity, bounds, equivalence, sign convention",
  "hydraulic_conductivity()", "11 soils × 15 h = 165 rows",
  "K(0)=Ks, monotonicity, bounds [0,Ks], Mualem constraint, validation errors",
  "fit_swrc()",               "11 soils × 12 h obs",
  "convergence, parameter recovery < 10 %, fitted-vs-observed visualisation",
  "Performance regressions",  "1 M rows (swrc, hc); 100 pedons seq + par",
  "< 5 s / < 10 s / < 30 s; parallel speedup > 1.2× on Unix"
) |>
  kable(caption = "Test coverage summary")
```
